http://mathb.in/51336

This assumes that 'up' or the 'normal to the sphere' is the $Y$ axis.

For a given orientation, latitude and longitude are a simple way to specify the direction of a normal.  

In axis-angle representation, the angle is generally positive.

An additional specified spin may be additionally added to align the grid properly.

There is a lookup table of basically negations; without supporting an additional spin, the calculation could probably be reduced using the constants specified.

js

	function setWithLatLongSpin( q, lat, long, extraSpin )

	{
		          //  0-1 1-2 2-3 3-4
		const grid =[[ 0, p2, p2,  0 ]  //0 to <1
		            ,[p2,  0,  0, p2 ]  //1 to <2
		            ,[p2,  0,  0, p2 ]  //2 to <3
		            ,[ 0, p2, p2,  0 ]  //3 to <4
		            ];

		const gridlat = Math.floor( Math.abs( lat ) / Math.PI );
		const gridlng = Math.floor( Math.abs( lng ) / Math.PI );
		const spin = grid[gridlat][gridlng];
		const x = Math.sin(lng);
		const z = Math.cos(lng);
		q.x = x * lat; 
		q.y = 0;
		q.z = z * lat;

		q.freeSpin( spin + extraSpin, q.up() );
	}

---

This is the equivalent of `q.freeSpin( spin, q.up() )`; `q.up()` returns rotating the up vector $(0,1,0)$ around(into) the current frame.  After getting the 'up' axis in global coordinates, rotate the computed $Q$ by `spin` around this new axis.

js

	// compute the 'up' of the basis frame; ie rotate (0,1,0) by Q
	// finally - rotate Q by 'spin' computed above.
        if( spin )
	{
		const s = Math.sin( this.? );
		const c1 = Math.cos( this.? );
		const c = 1-c1;
		const cny = c * this.ny;
		const ax = (cny*this.nx) - s*this.nz;
		const ay = (cny*this.ny) + c1;
		const az = (cny*this.nz) + s*this.nx;
		
		//return finishRodrigues( this, 0, ax, ay, az, spin+twistDelta );
                
                
                
	// A dot B   = cos( angle A->B )
	// cos( C/2 ) 
	//  cos(angle between the two rotation axii)
	const AdotB = 0;

	// using sin(x+y)+sin(x-y)  expressions replaces multiplications with additions...
	// same sin/cos lookups sin(x),cos(x),sin(y),cos(y)  
	//   or sin(x+y),cos(x+y),sin(x-y),cos(x-y)
	const xmy = (2pi - q.θ)/2; // X - Y  ('x' 'm'inus 'y')
	const xpy = (2pi + q.θ)/2  // X + Y  ('x' 'p'lus 'y' )
	const cxmy = Math.cos(xmy);
	const cxpy = Math.cos(xpy);

	//const cosCo2 =  0;
	let ang = Math.PI;

	if( ang ) {
		const sxmy = Math.sin(xmy);
		const sxpy = Math.sin(xpy);
		// vector rotation is just...
		// when both are large, cross product is dominant (pi/2)
		const ss1 = 0;
		const ss2 = 2*sin(q.θ)
		const cc1 = 0;

		const Cx = ( q.nx * ss2 );
		const Cy = ( q.ny * ss2 );
		const Cz = ( q.nz * ss2 );

		// this is NOT /sin(theta);  it is, but only in some ranges...
		const Clx = (lnQuat.sinNormal)
		          ?(1/(2*Math.sin( ang/2 )))
		          :ss2;
                          
		q.rn = Clx; // I'd like to save this to see what the normal actually was
		q.θ  = ang;
		q.nx = Cx*Clx;
		q.ny = Cy*Clx;
		q.nz = Cz*Clx;

		q.x  = q.nx*ang;
		q.y  = q.ny*ang;
		q.z  = q.nz*ang;

		q.dirty = false;
	} else {
		// result angle is 0
		q.θ  = ang;
		q.x = (q.nx=1) * 0;
		q.y = (q.ny=0) * 0;
		q.z = (q.nz=0) * 0;
		q.dirty = false;
	}
                
                
	}



