http://mathb.in/45267
http://mathb.in/64405 (adds alternative rodrigues)
http://mathb.in/64565 (Add get roll/pitch/yaw)
http://mathb.in/64566 (Remove 1-norm notations)
http://mathb.in/70972 (re-ordered)
---
Notation:
 
 - a vector : $[a,b,c,...]$
 - unit normal of a vector: $ ||[x,y,z]|| = \sqrt { x^2+y^2+z^2 } $
   - or for: $ [v_1,v_2,v_3,..] $, the 2-norm is: $ \sqrt { \sum_{i=1}^n v_i^2 } $

---
## AAA to axis-angle

Given $A$ as angle-angle-angle vector $[{X},{Y} ,{Z}] $ 

Axis-Angle: $ [ \frac {A} {||A||}, ||{A}|| ] $

---
## axis-angle to AAA

given axis-angle:
$ A_a =[ {x},{y},{z},{{\theta}}]  $, $A_r=[x,y,z]$, and $A_\theta = \theta$.

Angle angle angle: $   \frac  {{A_\theta}A_r} {||A_r||} $

or assuming that the axis is already a unit vector

Angle angle angle: $   {{A_\theta}A_r}$

---


## AAA To Quaternion

Given $A$ as angle-angle-angle vector $[{X},{Y} ,{Z}]$

Quaternion: [$\cos(\frac {||A||} {2} ),  \sin ( \frac {||A||} {2} ) \frac A {|||A||} $]


## Quaternion to AAA

Given $A$ as angle-angle-angle vector $[{X},{Y} ,{Z}]$

Quaternion: [$\cos(\frac {||A||} {2} ),  \sin ( \frac {||A||} {2} ) \frac A {|||A||} $]


---
## Conversion from Normal Vector(Y)

Representation of a normal as a rotation, this assumes that the vector $(0,1,0)$ is 'up'.  If some other axle is considered primary, the coordinates can be simply swapped.

This assume a unit vector in the direction of the normal
$N = [ X, Y, Z ]$


The rotation axis is the $X$,$Z$ coordinate(or $Y$,$X$ if $Z$ is 'up', or $Y$,$Z$ if $X$ is 'up')...
$   \alpha = |X| + |Z|$

If $\alpha = 0$ then result: $[0,0,0]$
else result: $ [      cos^{-1}( \frac {Y} {||N||} ) \frac X \alpha, 	0, cos^{-1}( \frac Y {||N||}) \frac Z \alpha ] $

## Align Normal using Basis
The default tangent and bi-tangent of rotations which only have their normal set, results in tangents and bi-tangents that are irregular.  Alternatively build a basis matrix, and convert from basis using the above mentioned method. 
Compute the normal of the above, and the matrix to convert...
And then use the basis to log quaternion conversion...

$\begin{matrix}    R =  [      cos^{-1}( \frac Y {||N||} ) \frac X \alpha, 	0, cos^{-1}( \frac Y {||N||} ) \frac Z \alpha ] \\
	E = \sqrt { R_Z^2+R_X^2 } \\
\begin{bmatrix}
(Y  \frac {-R_X }{ E })&X&\frac {R_Z}{E} \\
 (Z  \frac {R_Z}{E})-(X \cdot \frac {-R_X }{ E } )&Y&0\\
 (-(Y  \frac {R_Z}{E} )&Z&\frac {-R_X }{ E }
\end{bmatrix}
\end{matrix}
$


## Align Normal Directly

Or This is the direct computation to result with a log quaternion; compute the above result vector and then...


This is the angle $angle = cos^{-1} ( ( Y + 1 ) * ( 1 - {\frac X E} ) / 2 - 1 ) $


$\begin{matrix}
        yz = sin \beta X \\
	xz = ( 2 - (1-\cos \beta) (X^2 + Z^2) )  {\frac Z E}\\
	xy = sin \beta  X {\frac Z E}   + sin \beta  Z  (1-{\frac X E})
\end{matrix} \\
 \begin{bmatrix}
                            \frac{yz}{\sqrt(yz^2 + xz^2 + xy^2)} &
							\frac{xz}{\sqrt(yz^2 + xz^2 + xy^2)} &
							\frac{xy}{\sqrt(yz^2 + xz^2 + xy^2)}
\end{bmatrix}                            
$

and finally compute the resulting log quaternion.

Result : $\begin{bmatrix}
  \frac{n_x \cdot angle}{|n_x| + |n_y| + |n_z| }&
  \frac{n_y \cdot angle}{|n_x| + |n_y| + |n_z| }&
  \frac{n_z \cdot angle}{|n_x| + |n_y| + |n_z| }
\end{bmatrix} $

---

## Rodriguez Rotation

The input for this is a lnQuat, and an axis to rotate the rotation around and an angle around the axis to rotate.

Normalized axis of rotation...

given: 

- ${A} = [X_1, Y_1 ,Z_1] $   
  - Spin Angle $\alpha = ||A||$
  - axis $A_x$ = $A/||A||$
- ${B} = [X_2, Y_2 ,Z_2] $  
  - Spin Angle $\beta = ||B||$
  - axis $B_x$ = $B/||B||$.

Rotation angle: $\cos \frac{\gamma}{2} = \cos \alpha   \cos \beta  + \sin \alpha    \sin \beta  (  A_x \cdot B_x )$

get the result angle: $ \gamma = 2 \cos^{-1}( \cos \frac {\gamma}{2} )$


Compute axis : $C$ = $  \sin \beta  \cos \alpha  (A_x)  + \sin \alpha  \cos \beta (B_x) + \sin \alpha \sin \beta  ( A_x \cdot B_x ) $

result : $\frac {\gamma C} {||C||}$

## Rodrigues Alternative 

$th$ = angle to rotate around axis

$a = [ax,ay,az]$ - axis to rotate around

$q_{\theta}$ - rotation around axis to rotate

$ q = [qx,qy,qz]$ - axis to rotate

${ang} = 2 \cdot \arccos  ( \frac { ({{qx} \cdot {ax}+{qy}\cdot{ay}+{qz}\cdot{az})}\cdot({\cos {\frac {{th} + {q_\theta}} 2}}-{\cos {\frac {{th} - {q_\theta}} 2}}) + ( {\cos {\frac {{th} - {q_\theta}} 2}} + {\cos {\frac {{th} + {q_\theta}} 2}} ) } 2   )$

${s_1} = {\sin {\frac {{th} - {q\theta}} 2}} + {\sin {\frac {{th} + {q\theta}} 2}} $

${s_2} = {\sin {\frac {{th} + {q\theta}} 2}} - {\sin {\frac {{th} - {q\theta}} 2}} $

${c_1} = {\cos {\frac {{th} - {q\theta}} 2}} - {\cos {\frac {{th} + {q\theta}} 2}} $

${axisTemp} = {[{a}] \times [{q}]}\cdot {c_1} + {a}\cdot{s_1} + {q}\cdot{s_2} $

${axis} = \frac {axisTemp} { ||{axisTemp}|| } $

Rodrigues Original Axis calc : $\frac {axisTemp} {2 \cdot \sin  \frac {ang} 2 } $



## AAA to Matrix

given: ${A} = [X, Y ,Z] $

convert to angle-axis  $ \theta = ||A|| $, and $[x,y,z]$ = $\begin{bmatrix} \frac {X} {||A||} & \frac {Y} {||A||} &  \frac {Z} {||A||} \end{bmatrix} $


Compute some partial expressions:$\begin{matrix}
   x_y = xy(1-\cos \theta) &   w_x = x\sin \theta  &  x_x = xx(1-\cos \theta) \\
   y_z = yz(1-\cos \theta) &    w_y = y\sin \theta &  y_y = yy(1-\cos \theta) \\
   x_z = xz(1-\cos \theta) &    w_z = z\sin \theta &  z_z = zz(1-\cos \theta) 
\end{matrix}\\$


Compute the resulting matrix: $\begin{bmatrix}
\cos \theta+x_x &  x_y + w_z      & w_y + x_z \\
 w_z + x_y      & \cos \theta+y_y & y_z - w_x \\
 x_z - w_y      &  w_x + y_z      & \cos \theta+z_z
\end{bmatrix} $
 

Expanded : $    \begin{bmatrix}
\cos \theta+x^2 (1-\cos \theta)     &   xy(1-\cos \theta) + z\sin \theta & y\sin \theta + xz(1-\cos \theta) \\
 z\sin \theta + xy(1-\cos \theta)   & \cos \theta+y^2 (1-\cos \theta)    & yz(1-\cos \theta) - x\sin \theta \\
xz(1-\cos \theta) - y\sin \theta    &  x\sin \theta + yz(1-\cos \theta)  & \cos \theta+z^{2}(1-\cos \theta)
\end{bmatrix} $


---

These assume that $(0,1,0)$ is 'up' and $(1,0,0)$ is right and $(0,0,1)$ is forward.
Inputs are $(q_x,q_y,q_z)$ which determine the unit normal rotation axis, and $q_{\theta}$ which is the angle around said axis.

## Get Roll

from AAA, calculating the 'right' of the matrix, and using the Y coordinate of the right access gives the local roll.

effective range from -180 to 180


${roll} = \arcsin ({ \sin q_{\theta} \cdot q_z  + (1- c1) \cdot q_x \cdot q_y } )$


## Get Yaw

From AAA, use, the 'right' frame, and using the Z and X values to calculate the yaw.  This is assuming that
the frame generally is oriented such that roll is near zero.

$ s = \sin q_θ$

$ c_1 = \cos q_θ$

$ r_x = \cos q_θ      + (1- \cos q_θ) \cdot q_x \cdot q_x $

$ {principal} = \arcsin( -s \cdot q_y  + (1- c_1)\cdot q_x \cdot q_z)$

    if( r_x < 0 ) return principal;
    if( principal < 0 ) {
        result= -π-principal
    } else {
        result= π-principal
    }                           	
	


## Get Pitch

Calculating the Y of the forward vector yields the pitch; effective range from -90 to 90

$ {pitch} = \arcsin ( \sin  q_θ \cdot q_x + (\cos q_θ-1) \cdot q_z \cdot q_y ) $

